Random Forests for Automatic Matching Bullet-Lands

### 2015-12-21 imaging-paper.Rnw
removing all specific formatting

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\journal{Journal of Forensic Science International}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
%\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[USenglish]{babel}
\usepackage[protrusion=true,expansion=true]{microtype} % Better typography
\usepackage{amsmath,amsfonts,amsthm} % Math packages
\usepackage[svgnames]{xcolor} % Enabling colors by their 'svgnames'
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{fix-cm}	 % Custom font sizes - used for the initial letter in the document
\usepackage{float}
\usepackage{subcaption}

@@

\begin{document}

\begin{frontmatter}

\title{Random Forests for Automatic Matching Bullet-Lands}
\tnotetext[mytitlenote]{Resources for this manuscript, including R code to reproduce any of the models, tables and figures, can be found at  \url{https://github.com/erichare/imaging-paper}.}

%% Group authors per affiliation:
\author{Eric Hare, Heike Hofmann, Alicia Carriquiry}
\address{Department of Statistics and Statistical Laboratory, Iowa State University}
%\fntext[myfootnote]{Since 1880.}

%% or include affiliations in footnotes:
%\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
%\ead[url]{www.elsevier.com}

%\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{erichare@iastate.edu}

% \address[mymainaddress]{Department of Statistics and Statistical Laboratory, Iowa State University}
% \address[mysecondaryaddress]{secondary address -- CS?}

@@

\begin{keyword}
\hh{XXX not the right keywords yet} 3d Topological Surface Measurement, Data visualization, Machine Learning, Feature importance.
\end{keyword}

\end{frontmatter}

\linenumbers

@@

\hh{
Todo list for journal shape-up: there is a lot of good material, we mainly need to focus on what story to tell and make that a solid argument.
\begin{enumerate}
\item take out a few things: no circular fit, no figures 9 and 10 
\item do we need the web app?
\item out: lots of the figures can go into a supplement, together with a discussion. In the paper we then have a summary of what can be seen, so that we can focus on the results. 
\item add: formula for loess fit and its s.e.
\item add: comparison of error rates to those of FEs in Hamby study.
\item change: emphasis in varying cross section. once a stable region is found, how little change is happening? 
\item add: Why do we take the lowest stable cross section?
\end{enumerate}
}

@@

\section{Introduction}

% Rough outline:
% \begin{enumerate}
% \item DONE Description of the problem: Standard operating practice, 2009 NAS report
% \item DONE Description of the Results
% \item DONE Walkthrough of the process
% \item \hh{XXX Alternative approaches: IBIS database and the NRC report\citep{nap:2008} against it\footnote{Fox report on the NRC report \url{http://goo.gl/XurSVl}.
% NRC report: \url{http://www.nap.edu/catalog/12162/ballistic-imaging}} }
% \end{enumerate}

\hh{XXX Alternative approach by \citet{roberge:2006}: using correlation to identify matches between pairs of bullets using a correlation measure, that `could be automated'. No algorithm implemented/available, no error rates of automated matching are given XXX}

Firearm examination is a forensic tool used to determine whether two bullets were fired from the same gun barrel. This process has broad applicability in terms of the conviction of criminals in the United States criminal justice system. Ballistics identification has long been considered an accepted and reliable procedure, but in the past ten years has undergone more significant scrutiny. In 2005, in \emph{United States vs. Green}, the court ruled that the forensic expert could not confirm that the bullet casings came from a specific weapon with certainty, but could merely ``describe" other casings which are similar. Further court cases in the late 2000s expressed caution about the use of ballistics identification evidence~\citep{giannelli:2011}.

In 2009, the National Academy of Sciences published a report questioning the scientific validity of many forensic methods including firearm examination. ~\citep{NAS:2009}. The report concluded that minimal research had been done to explore the reliability of these ballistic identification methods. In fact, the report states that forensic evidence ``including, for example, bite marks and firearm and toolmark identification is introduced in criminal trials without any meaningful scientific validation, determination of error rates, or reliability testing to explain the limits of the discipline."

@@

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:residuala}Residual structure of a surface cross section at $x = 1.5625$ (bottom of the bullet).}
<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='.7\\linewidth'>>=
qplot(data=subset(dbr111, x <= 1.5625), y, value-ypred, #colour=factor(x),
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  theme_bw() + 
  geom_hline(yintercept = 0, colour="grey50") +
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:residualb} Residual structure of a surface cross section at $x = 100.00$}
<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='.7\\linewidth'>>=
#qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred,
qplot(data=subset(dbr111, x == 100), y, value-ypred,  #colour=factor(x), 
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  geom_hline(yintercept = 0, colour="grey50") +
  theme_bw() + 
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}
\caption{\label{fig:residual} Residual structure of circular fits at two different cross sections. Both residual plots show systematic structures, indicating that a circular fit is not entirely appropriate.}
\end{figure}

@@

\begin{figure}[hbtp]
  \centering
<<circlefits, echo=FALSE, fig.width=7, fig.height=7, out.width='\\linewidth', warning=FALSE>>=
qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
      geom="line", size=I(.75), colour=I("grey70")) + 
  facet_wrap(~part, ncol=2) + scale_colour_brewer(palette="Paired") +
  theme_bw() + theme(legend.position="bottom") + 
  geom_line(aes(y, resid, group = x), colour="black", size=.75, alpha=0.5,
             data = filter(db1, part==5, x == 100)[,c("y", "resid", "x")]) +
  ylab("Residuals from circular fit") +
  ggtitle("Bullet 1-5 in black")
@
\caption{Circle fit for each cross section of bullet 2, with bullet 1-5 overlaid.\label{fig:circlefits} Bullet 1-5 matches best with the bullet 2-1.}
\end{figure}

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:loess_step1}Step 1 the loess fit algorithm. For a fixed cross section ($x$ held constant at 243.75 $\mu m$) of surface measurements for bullet 1-5 are plotted across different values of $y$.}
<<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE,  message=FALSE>>=
br111_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
br121_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-1.x3p", sep = "/"))
br122_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-2.x3p", sep = "/"))
br123_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-3.x3p", sep = "/"))
br124_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-4.x3p", sep = "/"))
br125_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-5.x3p", sep = "/"))
br126_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-6.x3p", sep = "/"))

p1 <- qplot(y, value, data = br111_bullet, size=I(1)) + theme_bw() + ylab("Surface measurements")
p1
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:loess_step3}Step 2 of the loess fit algorithm. The  height values are smoothed twice with a smoothing factor of 35. The orange rectangle shows an example of the smoothing window. Grooves and peaks are detected, if they are not within the same window.}
<<loess_step3, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE,  message=FALSE>>=
subdata <- br111_bullet
smoothfactor <- 35

@@

p3
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:loess_step47}Steps 3 -- 6 of the loess fitting algorithm. After smoothing the $y$ values left and right minima are detected (marked by vertical lines, red indicating peaks and blue indicating valleys). Values outside these boundaries are removed (shown in grey)}
<<loess_step47, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE,  message=FALSE>>=
    lengthdiff <- length(subdata$value) - length(smoothed_truefalse)

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:loess_fit} Loess fit for bullet 1-5.}
<<loess_fit, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE,  message=FALSE>>=

@@

my.loess$fitted + ylab("Surface measurements")
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:loess_resid} Residuals of loess fit for bullet 1-5.}
<<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE>>=
my.loess$resid + ylab("Residuals of loess Fit")
@
\end{subfigure}
\caption{Fit and residuals of a loess fit to bullet 1-5 (Barrel 1).}
\end{figure}

\begin{figure}[hbtp]
  \centering
<<loessfits, echo=FALSE, fig.width=7, fig.height=7, out.width='\\linewidth', warning=FALSE>>=
br121.groove <- get_grooves(br121_bullet)
br122.groove <- get_grooves(br122_bullet)
br123.groove <- get_grooves(br123_bullet)
br124.groove <- get_grooves(br124_bullet)
br125.groove <- get_grooves(br125_bullet)
br126.groove <- get_grooves(br126_bullet)

@@

qplot(y, resid, data=melt.df, #colour=factor(x), 
      geom="line", size=I(.75), colour=I("grey70")) + 
  facet_wrap(~part, ncol=2) + scale_colour_brewer(palette="Paired") +
  theme_bw() + theme(legend.position="bottom") + 
  geom_line(data = actual.df, aes(y, resid), colour="black", size=.75, alpha=0.5) + ggtitle("Bullet 1-5 in black") + ylab("Residuals of loess Fit")
@    
\caption{\label{fig:loessfits} Residuals of a loess fit to bullets 1 and 2 from barrel 1. Bullet 1-5 is shown in black. }
\end{figure}

@@

lines <- matches %>% group_by(lineid) %>% summarise(
  meany = mean(y, na.rm=T),
  miny = min(y, na.rm=T),
  maxy = max(y, na.rm=T) + 1.5625,
  match = isMatch(lineid, type),
  type = type1[1]
)
lines <- subset(lines, lineid != 0)
@
\begin{figure}[hbtp]
\centering
\begin{subfigure}[t]{\linewidth}\centering
\caption{Cross section at $x = 100$.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.7\\textwidth'>>=
qplot(y, resid, group=bullet, colour=bullet, data=lof, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(.9, .2),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}\centering
    \caption{Adjusted (bullet 1-5 horizontally shifted) cross sections at $x = 100$.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.7\\textwidth'>>=
# qplot(y, resid, group=bullet, colour=bullet, data=lofX, 
#       geom="line", alpha=I(0.6)) + 
#   scale_colour_brewer(palette="Set1") + theme_bw() +
#   theme(legend.position="none")
qplot(y, resid, group=bullet, colour=bullet, data=lofY, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position="none") + ylab("Residuals of loess fit")
@
    }
\end{subfigure}
\caption{\label{fig:cross100}Lines showing surface cross sections at $x = 100$. A horizontal shift of the values of bullet 1-5 to the right shows the similarity of the striation marks.}
\end{figure}

@@

\begin{figure}[hbtp]
  \centering
<<ccf, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '.7\\linewidth'>>=
lag <- ccf$lag[which.max(ccf$acf)]
qplot(x=ccf$lag, xend=ccf$lag, y = 0, yend= ccf$acf, geom="segment", colour = I("grey50")) + 
  theme_bw() + ylab("Correlation") + xlab("Lag k") +
  geom_segment(aes(x = lag, xend = lag, y = 0, yend =  max(ccf$acf)), colour="black") + 
  scale_x_continuous(breaks = c(-100, -50, lag, 0, 50, 100),
                     minor_breaks = c(-75,-25,25, 75))
@
\caption{\label{fig:ccf}Cross-correlation between the two cross sections shown in Figure~\ref{fig:crosscut} at lags between -100 and 100. At a lag of -10 the correlation peaks, indicating the largest amount of agreement between the cross sections. Figure~\ref{fig:crosscutX} shows the lag-shifted cross sections.}
\end{figure}

\subsection{Varying cross sections}
The cross section at which a comparison is done matters -- as cross sections are further apart, more pronounced differences between the cross sections show up. This poses both a caveat to matching attempts as well as an opportunity \hh{for quality control}: on the one hand we have to be aware of which cross section level was used in a matching, but on the other hand we can use multiple cross sections of the same bullet for an initial assessment of its quality. Both of these situations are assessed and discussed in more detail below: 
Figure~\ref{fig:crosscuts} shows a sequence of cross sections for bullet 1-5 (barrel 1) that are taken 50$\mu m$ apart, between 150$\mu m$  and 400$\mu m$. These are compared to a cross section at 100$\mu m$. Initially this comparison constitutes an almost perfect match between the two cross sections. However, the match quickly deteriorates with increasing distance to non-matches at a cross section of $x = 300\mu m$ or higher.  Only if the cross section is within 150$\mu m$ do we get a good visual match even when we know that the same bullet surface is being used. 
Given that we have to expect some variation in nominally the same cross section levels due to manual alignments in microscopes, we should take cross section levels into account in the automatic matching routine by evaluating matches at several cross sections. 
\begin{figure}[hbtp]
  \centering
<<crosscuts-vary, echo=FALSE, fig.width = 7, fig.height = 12, out.width = '.7\\linewidth', cache=TRUE, warning = FALSE>>=
paths <- file.path(datadir, dir(datadir))
crosscuts <- seq(100, 400, by = 50)
lof <- processBullets(paths[5], x = crosscuts)
lof$bullet <- paste(lof$bullet, lof$x)

@@

\begin{figure}[hbtp]
  \centering
<<crosscuts-vary-b31, echo=FALSE, fig.width = 7, fig.height = 10, out.width = '.7\\linewidth', cache=TRUE, warning = FALSE>>=
paths <- file.path(datadir, dir(datadir))
crosscuts <- seq(100,225, by=25)
lof <- processBullets(paths[37], x = crosscuts)
lof$bullet <- paste(lof$bullet, lof$x)

@@

\begin{figure}[hbtp]
\centering
    \begin{subfigure}[t]{\linewidth}\centering
\caption{Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}}{%
<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.7\\textwidth', warning=FALSE>>=
qplot(data=lofX, y, resid, colour=I("grey30"), size=I(0.75), geom="line", group=bullet) +
  geom_line(aes(y=l30), colour="grey70", size=0.5) +
  facet_grid(bullet~.) +
#  scale_colour_manual("", values=cols) +
  theme_bw() + ylab("") + 
  theme(legend.position="none")
@
    }
\end{subfigure}
\begin{subfigure}[t]{\linewidth}\centering
\caption{Using a rolling median peaks and grooves are identified for each cross section. \label{fig:smoothcutb}}{%
<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.7\\textwidth', warning=FALSE>>=
#images <- file.path(datadir, dir(datadir))
  lof <- processBullets(paths = images[c(5,7)], x = 100)
  lof <- bulletSmooth(lof)
  bAlign = bulletAlign(lof)
  lofX <- bAlign$bullet

@@

\begin{figure}[hbtp]
  \centering
<<smoothfac1, echo=FALSE, fig.width=7, fig.height=5.2, out.width='.7\\linewidth', warning=FALSE>>=
br111.groove <- get_grooves(br111_bullet)
br111.loess <- fit_loess(br111_bullet, br111.groove)

@@

\begin{figure}[hbtp]
  \centering
<<barrel2, echo=FALSE, fig.width = 7, fig.height=12, out.width='.7\\linewidth', warning=FALSE>>=

@@

\section*{References}

\bibliography{references}

>>>>

\documentclass[11pt]{article}
\usepackage[margin = 1in]{geometry}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]
\usepackage{setspace}
\doublespacing

\usepackage[USenglish]{babel}
\usepackage[protrusion=true,expansion=true]{microtype} % Better typography
\usepackage{amsmath,amsfonts,amsthm} % Math packages
\usepackage[svgnames]{xcolor} % Enabling colors by their 'svgnames'
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{fix-cm}	 % Custom font sizes - used for the initial letter in the document
\usepackage{natbib}
\usepackage{float}
\usepackage{subcaption}

@@

\title{Random Forests for Automatic Matching Bullet-Lands}
\author{Eric Hare{\footnote{Corresponding author: erichare@iastate.edu}}, Heike Hofmann, Alicia Carriquiry\\
Department of Statistics and Statistical Laboratory, Iowa State University}

\begin{document}
\maketitle

@@

\paragraph{Keywords:} \hh{XXX not the right keywords yet} 3d Topological Surface Measurement, Data visualization, Machine Learning, Feature importance.

\newpage
\linenumbers

@@

@@

\section{Introduction}
\hh{
Todo list for journal shape-up: there is a lot of good material, we mainly need to focus on what story to tell and make that a solid argument.
\begin{enumerate}
\item take out a few things: no circular fit, no figures 9 and 10 
\item do we need the web app?
\item out: lots of the figures can go into a supplement, together with a discussion. In the paper we then have a summary of what can be seen, so that we can focus on the results. 
\item add: formula for loess fit and its s.e.
\item add: comparison of error rates to those of FEs in Hamby study.
\item change: emphasis in varying cross section. once a stable region is found, how little change is happening? 
\item add: Why do we take the lowest stable cross section?
\end{enumerate}
}
Firearm examination is a forensic tool used to determine whether two bullets were fired from the same gun barrel. This process has broad applicability in terms of the conviction  in the United States criminal justice system. Ballistics identification has long been considered an accepted and reliable procedure, but in the past ten years has undergone more significant scrutiny. In 2005, in \emph{United States vs. Green}, the court ruled that the forensic expert could not confirm that the bullet casings came from a specific weapon with certainty, but could merely ``describe" other casings which are similar. Further court cases in the late 2000s expressed caution about the use of ballistics identification evidence~\citep{giannelli:2011}.

In 2009, the National Academy of Sciences published a report questioning the scientific validity of many forensic methods including firearm examination. ~\citep{NAS:2009}. The report states that ``[m]uch forensic evidence -- including, for example, bite marks and firearm and toolmark identification is introduced in criminal trials without any meaningful scientific validation, determination of error rates, or reliability testing to explain the limits of the discipline."

@@

\hh{XXX More detailed discussion of automatic approaches and database discussion XXX}
\hh{XXX Alternative approach by \citet{roberge:2006}: using correlation to identify matches between pairs of bullets using a correlation measure, that `could be automated'. No algorithm implemented/available, no error rates of automated matching are given XXX}

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:residuala}Residual structure of a surface cross section at $x = 1.5625$ (bottom of the bullet).}
<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='.5\\linewidth'>>=
qplot(data=subset(dbr111, x <= 1.5625), y, value-ypred, #colour=factor(x),
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  theme_bw() + 
  geom_hline(yintercept = 0, colour="grey50") +
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:residualb} Residual structure of a surface cross section at $x = 100.00$}
<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='.5\\linewidth'>>=
#qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred,
qplot(data=subset(dbr111, x == 100), y, value-ypred,  #colour=factor(x), 
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  geom_hline(yintercept = 0, colour="grey50") +
  theme_bw() + 
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}
\caption{\label{fig:residual} Residual structure of circular fits at two different cross sections. Both residual plots show systematic structures, indicating that a circular fit is not entirely appropriate.}
\end{figure}

@@

\begin{figure}[hbtp]
  \centering
<<circlefits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\linewidth', warning=FALSE>>=
qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
      geom="line", size=I(.75), colour=I("grey70")) + 
  facet_wrap(~part, ncol=3) + scale_colour_brewer(palette="Paired") +
  theme_bw() + theme(legend.position="bottom") + 
  geom_line(aes(y, resid, group = x), colour="black", size=.75, alpha=0.5,
             data = filter(db1, part==5, x == 100)[,c("y", "resid", "x")]) +
  ylab("Residuals from circular fit") +
  ggtitle("Bullet 1-5 in black")
@
\caption{Circle fit for each cross section of bullet 2, with bullet 1-5 overlaid.\label{fig:circlefits} Bullet 1-5 matches best with the bullet 2-1.}
\end{figure}

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:loess_step1}Step 1 the loess fit algorithm. For a fixed cross section ($x$ held constant at 243.75 $\mu m$) of surface measurements for bullet 1-5 are plotted across different values of $y$.}
<<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\linewidth', warning=FALSE,  message=FALSE>>=
br111_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
br121_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-1.x3p", sep = "/"))
br122_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-2.x3p", sep = "/"))
br123_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-3.x3p", sep = "/"))
br124_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-4.x3p", sep = "/"))
br125_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-5.x3p", sep = "/"))
br126_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-6.x3p", sep = "/"))

p1 <- qplot(y, value, data = br111_bullet, size=I(1)) + theme_bw() + ylab("Surface measurements")
p1
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:loess_step3}Step 2 of the loess fit algorithm. The  height values are smoothed twice with a smoothing factor of 35. The orange rectangle shows an example of the smoothing window. Grooves and peaks are detected, if they are not within the same window.}
<<loess_step3, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\linewidth', warning=FALSE,  message=FALSE>>=
subdata <- br111_bullet
smoothfactor <- 35

@@

p3
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:loess_step47}Steps 3 -- 6 of the loess fitting algorithm. After smoothing the $y$ values left and right minima are detected (marked by vertical lines, red indicating peaks and blue indicating valleys). Values outside these boundaries are removed (shown in grey)}
<<loess_step47, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\linewidth', warning=FALSE,  message=FALSE>>=
    lengthdiff <- length(subdata$value) - length(smoothed_truefalse)

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:loess_fit} Loess fit for bullet 1-5.}
<<loess_fit, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\linewidth', warning=FALSE,  message=FALSE>>=

@@

my.loess$fitted + ylab("Surface measurements")
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:loess_resid} Residuals of loess fit for bullet 1-5.}
<<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\linewidth', warning=FALSE>>=
my.loess$resid + ylab("Residuals of loess Fit")
@
\end{subfigure}
\caption{Fit and residuals of a loess fit to bullet 1-5 (Barrel 1).}
\end{figure}

\begin{figure}[hbtp]
  \centering
<<loessfits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\linewidth', warning=FALSE>>=
br121.groove <- get_grooves(br121_bullet)
br122.groove <- get_grooves(br122_bullet)
br123.groove <- get_grooves(br123_bullet)
br124.groove <- get_grooves(br124_bullet)
br125.groove <- get_grooves(br125_bullet)
br126.groove <- get_grooves(br126_bullet)

@@

qplot(y, resid, data=melt.df, #colour=factor(x), 
      geom="line", size=I(.75), colour=I("grey70")) + 
  facet_wrap(~part, ncol=3) + scale_colour_brewer(palette="Paired") +
  theme_bw() + theme(legend.position="bottom") + 
  geom_line(data = actual.df, aes(y, resid), colour="black", size=.75, alpha=0.5) + ggtitle("Bullet 1-5 in black") + ylab("Residuals of loess Fit")
@    
\caption{\label{fig:loessfits} Residuals of a loess fit to bullets 1 and 2 from barrel 1. Bullet 1-5 is shown in black. }
\end{figure}

@@

lines <- matches %>% group_by(lineid) %>% summarise(
  meany = mean(y, na.rm=T),
  miny = min(y, na.rm=T),
  maxy = max(y, na.rm=T) + 1.5625,
  match = isMatch(lineid, type),
  type = type1[1]
)
lines <- subset(lines, lineid != 0)
@
\begin{figure}[hbtp]
\centering
\begin{subfigure}[t]{\linewidth}\centering
\caption{Cross section at $x = 100$.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.5\\textwidth'>>=
qplot(y, resid, group=bullet, colour=bullet, data=lof, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(.9, .2),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}\centering
    \caption{Adjusted (bullet 1-5 horizontally shifted) cross sections at $x = 100$.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.5\\textwidth'>>=
# qplot(y, resid, group=bullet, colour=bullet, data=lofX, 
#       geom="line", alpha=I(0.6)) + 
#   scale_colour_brewer(palette="Set1") + theme_bw() +
#   theme(legend.position="none")
qplot(y, resid, group=bullet, colour=bullet, data=lofY, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position="none") + ylab("Residuals of loess fit")
@
    }
\end{subfigure}
\caption{\label{fig:cross100}Lines showing surface cross sections at $x = 100$. A horizontal shift of the values of bullet 1-5 to the right shows the similarity of the striation marks.}
\end{figure}

@@

\begin{figure}[hbtp]
  \centering
<<ccf, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '.5\\linewidth'>>=
lag <- ccf$lag[which.max(ccf$acf)]
qplot(x=ccf$lag, xend=ccf$lag, y = 0, yend= ccf$acf, geom="segment", colour = I("grey50")) + 
  theme_bw() + ylab("Correlation") + xlab("Lag k") +
  geom_segment(aes(x = lag, xend = lag, y = 0, yend =  max(ccf$acf)), colour="black") + 
  scale_x_continuous(breaks = c(-100, -50, lag, 0, 50, 100),
                     minor_breaks = c(-75,-25,25, 75))
@
\caption{\label{fig:ccf}Cross-correlation between the two cross sections shown in Figure~\ref{fig:crosscut} at lags between -100 and 100. At a lag of -10 the correlation peaks, indicating the largest amount of agreement between the cross sections. Figure~\ref{fig:crosscutX} shows the lag-shifted cross sections.}
\end{figure}

\subsection{Varying cross sections}
The cross section at which a comparison is done matters -- as cross sections are further apart, more pronounced differences between the cross sections show up. This poses both a caveat to matching attempts as well as an opportunity \hh{for quality control}: on the one hand we have to be aware of which cross section level was used in a matching, but on the other hand we can use multiple cross sections of the same bullet for an initial assessment of its quality. Both of these situations are assessed and discussed in more detail below: 
Figure~\ref{fig:crosscuts} shows a sequence of cross sections for bullet 1-5 (barrel 1) that are taken 50$\mu m$ apart, between 150$\mu m$  and 400$\mu m$. These are compared to a cross section at 100$\mu m$. Initially this comparison constitutes an almost perfect match between the two cross sections. However, the match quickly deteriorates with increasing distance to non-matches at a cross section of $x = 300\mu m$ or higher.  Only if the cross section is within 150$\mu m$ do we get a good visual match even when we know that the same bullet surface is being used. 
Given that we have to expect some variation in nominally the same cross section levels due to manual alignments in microscopes, we should take cross section levels into account in the automatic matching routine by evaluating matches at several cross sections. 
\begin{figure}[hbtp]
  \centering
<<crosscuts-vary, echo=FALSE, fig.width = 7, fig.height = 12, out.width = '.5\\linewidth', cache=TRUE, warning = FALSE>>=
paths <- file.path(datadir, dir(datadir))
crosscuts <- seq(100, 400, by = 50)
lof <- processBullets(paths[5], x = crosscuts)
lof$bullet <- paste(lof$bullet, lof$x)

@@

\begin{figure}[hbtp]
  \centering
<<crosscuts-vary-b31, echo=FALSE, fig.width = 7, fig.height = 10, out.width = '.5\\linewidth', cache=TRUE, warning = FALSE>>=
paths <- file.path(datadir, dir(datadir))
crosscuts <- seq(100,225, by=25)
lof <- processBullets(paths[37], x = crosscuts)
lof$bullet <- paste(lof$bullet, lof$x)

@@

\begin{figure}[hbtp]
\centering
    \begin{subfigure}[t]{\linewidth}\centering
\caption{Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}}{%
<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE>>=
qplot(data=lofX, y, resid, colour=I("grey30"), size=I(0.75), geom="line", group=bullet) +
  geom_line(aes(y=l30), colour="grey70", size=0.5) +
  facet_grid(bullet~.) +
#  scale_colour_manual("", values=cols) +
  theme_bw() + ylab("") + 
  theme(legend.position="none")
@
    }
\end{subfigure}
\begin{subfigure}[t]{\linewidth}\centering
\caption{Using a rolling median peaks and grooves are identified for each cross section. \label{fig:smoothcutb}}{%
<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE>>=
#images <- file.path(datadir, dir(datadir))
  lof <- processBullets(paths = images[c(5,7)], x = 100)
  lof <- bulletSmooth(lof)
  bAlign = bulletAlign(lof)
  lofX <- bAlign$bullet

@@

\begin{figure}[hbtp]
  \centering
<<smoothfac1, echo=FALSE, fig.width=7, fig.height=5.2, out.width='.5\\linewidth', warning=FALSE>>=
br111.groove <- get_grooves(br111_bullet)
br111.loess <- fit_loess(br111_bullet, br111.groove)

@@

\begin{figure}[hbtp]
  \centering
<<barrel2, echo=FALSE, fig.width = 7, fig.height=12, out.width='.5\\linewidth', warning=FALSE>>=

@@

Resources for this manuscript, including R code to reproduce any of the models, tables and figures, can be found at  \url{https://github.com/erichare/imaging-paper}.
